R Notebook

### 2017-04-06 doc/RMD/MLG-distribution.Rmd
rename populations

```{r read_data}
dat <- read.genalex("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", ploidy = 1)
splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
dat

@@

```{r}
rerange <- function(x){
  minx <- min(x, na.rm = TRUE)
  maxx <- max(x, na.rm = TRUE)
  if (minx < 0)
    x <- x + abs(minx)
    maxx <- maxx + abs(minx)
  if (maxx > 1)
    x <- x/maxx
  return(x)
}
make_graph_list <- function(dat){ # dat is a genclone/snpclone object
  # w       <- pgen(dat, by_pop = FALSE) %>% rowSums() %>% exp() %>% setNames(paste0("MLG.", mll(dat)))
  # w       <- w[unique(names(w))]
  w       <- psex(dat, by_pop = FALSE, method = "multiple") %>% split(mll(dat))
  names(w)<- paste0("MLG.", names(w))
  datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
  crosses <- mlg.crosspop(dat, quiet = TRUE, df = TRUE) %>% tbl_df()
  adjmat  <- datmlg %*% t(datmlg) 
  cols    <- sort(colnames(adjmat))
  adjmat  <- adjmat[cols, cols]
  # Creating Graph
  g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
  V(g)$size   <- diag(adjmat)
  g           <- delete_vertices(g, degree(g) == 0)
  shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
  V(g)$weight <- 1 - shared_mlg/V(g)$size # fraction of private MLGs
  el          <- as_adj_edge_list(g)
  el          <- el[lengths(el) > 0]
  popgraphs <- setNames(vector(mode = "list", length = length(el) + 1), c(names(el), "total"))
  for (v in names(el)){
    idx  <- el[[v]]
    mlgs <- crosses %>%           # How to get all MLGs from a single population:
      filter(Population == v) %>%         # Grab only the population e and then
      select(MLG) %>%                     # remove everything but the MLGs to do an
      inner_join(crosses, by = "MLG") %>% # inner join of the original list and then
      filter(Population != v) %>%         # remove the query population to give
      arrange(Population)                 # the neigboring populations in order.
    MLGS <- as.character(mlgs$MLG)
    E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
    E(g)[idx]$width  <- as.integer(table(MLGS)[MLGS]) # size == n populations visited
    E(g)[idx]$weight <- 1 - map_dbl(w[MLGS], 2)       # weight == 1 - psex 2nd encounter
    popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
  }
  popgraphs[["total"]] <- g
  popgraphs
}

plot_mlg_graph <- function(g, glayout = NULL){
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- (1 - V(g)$weight) * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size)) +
    geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) +
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
      title = "Shared haplotypes across regions",
      fill = "Number of\nGenotypes",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nPer haplotype",
      caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
    ))
}

@@

good_layout <- read.table(
text = 
"   x                            y
AU  0.8090169943749470071737  5.877852522924730260812e-01
CA -1.0000000000000000000000  1.224646799147350002426e-16
CO -0.3090169943749480063744 -9.510565162951539752711e-01
FR -0.3090169943749470071737  9.510565162951539752711e-01
MI  0.0000000000000000000000  0.000000000000000000000e+00
MN  1.0000000000000000000000  0.000000000000000000000e+00
ND  0.3090169943749470071737 -9.510565162951539752711e-01
NE -0.8090169943749470071737 -5.877852522924730260812e-01
NY  0.8090169943749470071737 -5.877852522924730260812e-01
OR  0.3090169943749470071737  9.510565162951539752711e-01
WA -0.8090169943749470071737  5.877852522924730260812e-01"
) %>% as.matrix()

>>>>

```{r read_data}
dat <- read.genalex("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", ploidy = 1)
splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
pops <- levels(strata(dat)$Region)
levels(strata(dat)$Region) <- case_when(pops == "FR" ~ "France", 
                                        pops == "MX" ~ "Mexico", 
                                        pops == "AU" ~ "Australia", 
                                        TRUE ~ pops)
dat

@@

```{r}
rerange <- function(x){
  minx <- min(x, na.rm = TRUE)
  maxx <- max(x, na.rm = TRUE)
  if (minx < 0)
    x <- x + abs(minx)
    maxx <- maxx + abs(minx)
  if (maxx > 1)
    x <- x/maxx
  return(x)
}
make_graph_list <- function(dat){ # dat is a genclone/snpclone object
  # w       <- pgen(dat, by_pop = FALSE) %>% rowSums() %>% exp() %>% setNames(paste0("MLG.", mll(dat)))
  # w       <- w[unique(names(w))]
  w       <- psex(dat, by_pop = FALSE, method = "multiple") %>% split(mll(dat))
  names(w)<- paste0("MLG.", names(w))
  datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
  crosses <- mlg.crosspop(dat, quiet = TRUE, df = TRUE) %>% tbl_df()
  adjmat  <- datmlg %*% t(datmlg) 
  cols    <- sort(colnames(adjmat))
  adjmat  <- adjmat[cols, cols]
  # Creating Graph
  g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
  V(g)$size   <- diag(adjmat)
  g           <- delete_vertices(g, degree(g) == 0)
  shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
  V(g)$weight <- shared_mlg/V(g)$size # fraction of shared MLGs
  el          <- as_adj_edge_list(g)
  el          <- el[lengths(el) > 0]
  popgraphs <- setNames(vector(mode = "list", length = length(el) + 1), c(names(el), "total"))
  for (v in names(el)){
    idx  <- el[[v]]
    mlgs <- crosses %>%           # How to get all MLGs from a single population:
      filter(Population == v) %>%         # Grab only the population e and then
      select(MLG) %>%                     # remove everything but the MLGs to do an
      inner_join(crosses, by = "MLG") %>% # inner join of the original list and then
      filter(Population != v) %>%         # remove the query population to give
      arrange(Population)                 # the neigboring populations in order.
    MLGS <- as.character(mlgs$MLG)
    E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
    E(g)[idx]$width  <- as.integer(table(MLGS)[MLGS]) # size == n populations visited
    E(g)[idx]$weight <- 1 - map_dbl(w[MLGS], 2)       # weight == 1 - psex 2nd encounter
    popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
  }
  popgraphs[["total"]] <- g
  popgraphs
}

plot_mlg_graph <- function(g, glayout = NULL, label = TRUE){
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- V(g)$weight * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  outg <- ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
  outg + 
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
      title = "Shared haplotypes across regions",
      fill = "Number of\nhaplotypes",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nper haplotype",
      caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
    ))
}

@@

good_layout <- read.table(
text = 
"   x                            y
Australia  0.8090169943749470071737  5.877852522924730260812e-01
CA -1.0000000000000000000000  1.224646799147350002426e-16
CO -0.3090169943749480063744 -9.510565162951539752711e-01
France -0.3090169943749470071737  9.510565162951539752711e-01
MI  0.0000000000000000000000  0.000000000000000000000e+00
MN  1.0000000000000000000000  0.000000000000000000000e+00
ND  0.3090169943749470071737 -9.510565162951539752711e-01
NE -0.8090169943749470071737 -5.877852522924730260812e-01
NY  0.8090169943749470071737 -5.877852522924730260812e-01
OR  0.3090169943749470071737  9.510565162951539752711e-01
WA -0.8090169943749470071737  5.877852522924730260812e-01"
) %>% as.matrix()

alt_layout <- read.table(
text = 
"   x                            y
Australia  0.8090169943749470071737  5.877852522924730260812e-01
OR -1.0000000000000000000000  1.224646799147350002426e-16
CO -0.3090169943749480063744 -9.510565162951539752711e-01
WA -0.3090169943749470071737  9.510565162951539752711e-01
MI  0.0000000000000000000000  0.000000000000000000000e+00
MN  1.0000000000000000000000  0.000000000000000000000e+00
ND  0.3090169943749470071737 -9.510565162951539752711e-01
NE -0.8090169943749470071737 -5.877852522924730260812e-01
NY  0.8090169943749470071737 -5.877852522924730260812e-01
France  0.3090169943749470071737  9.510565162951539752711e-01
CA -0.8090169943749470071737  5.877852522924730260812e-01"
) %>% as.matrix()
alt_layout <- alt_layout[rownames(good_layout), ]
